Learning with misspecified models: the case of overestimation

### 2023-10-14 Paper/draft.tex
smm

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Learning with misspecified models: Overconfidence and Stereotypes},
  pdfauthor={Jimena Galindo},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=2.5cm]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

@@

\title{Learning with misspecified models: Overconfidence and
Stereotypes}
\author{Jimena Galindo}
\date{October 10, 2023}

\begin{document}
\maketitle
\begin{abstract}
TBW
\end{abstract}

@@

A growing body of literature in economics studies the mechanisms through
which economic agents come to hold incorrect beliefs about a
fundamental. Most of this literature focuses on the case in which agents
are passive observers of the world and incorrectly incorporate the
information they receive into their
beliefs\footnote{See \citet{benjamin2019} for a review of error in probabilistic thinking}.
However, in many situations, agents are active participants in the
process of generating information. In such cases, the information they
receive is affected by their actions and the way they process it can
affect their future behavior. I refer to this as an
\emph{endogenous information} setting. This sort of setting, has been
shown to explain pehnomena that cannot be rationalized with the
classical models of biased belief updating (e.g. \citet{Nyarko1991} and
\citet{Levy2022}). Settings with endogenous information and misspecified
learning also offer alternative explanations for why certain well known
biases might persist. I focus on a group of theories that aim to explain
the prevalence of overconfidence.

Overconfidence has been shown to be a prevalent bias in a variety of
settongs. \citet{Oster2013} shows that subjects who are at high risk of
having Huntingtons disease underestimate their probability of having the
disease and make retirement decisions as if they were not sick. Relating
to investment decisions, \citet{Hoffman2020} show that workers
overestimate their productivity in their current employment and are
unlikely to look for other oportunities. \citet{Camerer1999} find that
entrepreneurs are overconfident about the quiality of their enterprise,
which leads to excessive entry and early exit from markets. These are
only some examples of how holding an incorrect belief about a
fundamental can lead to costly decisions \footnote{Other examples 
of overconfidence are }.

I isolate the main forces from 3 different theories of misspecified
learning. Each of the theories rationalizes the prevalence of
overconfidence in frameworks where the agents are learning about an
ego-relevant parameter and an exogenously determined state. The
interaction between these two parameters together with an endogenous
information process give rise to two explanaitions for the presistence
of overconfidence. On one hand, even with bayesian updating, the agent
may fall into learning traps. Learning traps are misspecified beliefs
and optimal actions that induce signals that confirm the misspecified
belief. Once an agent finds themselves in such a situation, the belief
will be stable and even with a correctly speficified model of learning,
they will not be able to abandon their icorrect beliefs. On the other
hand a setting with two variables allows the agent to engage in
asymmetric updating with respect to good and bad news. When the outcome
depends on an ego-relevant parameter as well as a state, the agent can
attribute good news to a high value of the parameter while attributing
failures to a low value of the exogenous
parameter\footnote{This is in line with the well known theory 
in behavioral science of self-attribution bias of which \citet{kelley1980} provides a review.}.

\citet{Heidhues2018} shows that under certain assumptions on the output
function, learning traps can rationalize the existance of misspecified
beliefs when agents are dogmatic about the ego-relevant parameter. In a
more general model \citet{Ba2023} further analyzes this setting and
characterizes the stuations in which even agents who are not dogmatic
might become trapped. She allows initi dogmatic agents to consider
alternative theories and adopt whichever is a better explanation of the
observed data. They do so by performing Bayesian hypothesis tests that
compare the likelihood of having observed the outcome under alternative
paradigms\footnote{A 
similar mechanism is proposed by \citet{Schwarstein2021} in a setting with persuasion}.
Although this learning mechanism allows agents to avoid the learning
traps, there are still situations under which misspecified beliefs may
persist.

@@

\emph{The student must choose an effort level \(a\), which determines
her grade. For whatever the chosen effort level, the student must pay a
cost \(c(a) = \frac{1}{2}a^2\) and receives a grade of
\((\theta^*+a)\omega^*\). She repeats this process for infinitely many
periods. Assume also that the student's prior is such that
\(\mathbb{E}[\omega]= \omega^*\) and she is dogmatic about being of type
\(\hat{\theta}>\theta^*\).\footnote{The example is illustrated for an overconfident student but 
the results are symmetric for a dogmatic student who initially 
places probability one on some $\tilde{\theta}<\theta^*$.} Therefore,
the student's payoff in period \(t\) is given by }

@@

\emph{Under this specification, the myopic optimal effort level is
\(a_t^* = \omega^*\). Because the student does not know \(\omega^*\), he
will choose \(a_t = \mathbb{E}_t(\omega)\) where the expectation is
taken with respect to the student's belief at the beginning of period
\(t\). If she does not revise his effort choice for \(k\) periods, she
will receive an average utility of
\((\theta^*+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}\) but he was expecting
an average utility of
\((\hat{\theta}\theta+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}\). In
response, she will apply Bayes rule to update his beliefs about
\(\omega\) to get the posterior belief with
\(\mathbb{E}_{t+k}[\omega] = \frac{(\theta^{*} + \omega^{*})\omega^{*}}{\hat{\theta} + \omega^{*}}\)
which is lower than the initial belief. This will cause the student to
choose a lower effort at \(t+k\). As a result, she will again receive an
average utility that is lower than what he expected which will cause her
belief to drift further down. This process will continue until the
average utility equals her expected utility under the dogmatic belief
that assigns probability 1 to \(\hat{\theta}\). At that point, the
student will have reached a self-defeating equilibrium and she will
continue to choose sub-optimal effort forever. }

@@

A regression analysis of this same data reveals that indeed, the slopes
are different for the two types of news and that the difference is
statistically significant (\(p<0.01\)). The regression results are
reported in Table \ref{tab:regression}. columns 2 and 3 of the table
consider the model for each treatment separately. I find no significant
differences in the parameters across treatment. This is consistent with
the fact that the bias is not driven by the ego-relevance of
\(\theta\).\footnote{The ego-relevance of $\theta$ is the motivation behind the self-attribution bias, 
however, the bias itself is due to an incorrect updating procedure in response to good and bad news. The presence of 
asymmetric updating in the stereotype condition can be due to self-censoring of the subjects. In particular, subjects 
who initially underestimated the performance of the other participant might have realized their mistake and 
by overestimating it after good news. This is consistent with the data presented here as well as with 
the data observed in the transition matrix for stereotypes which is reported in the Appendix}

@@

More convincing evidence against the Switching model is the lack of
inverse or attenuated correlation between the change in effort and the
news differential. In particular, if the switcher were the underlying
model, I would expect overconfident subjects who start with an
overestimation of \(\theta\) to increase their effort in response to bad
news that were very surprising. This would flatten the slope of the
regression line for bad news relative to what a Bayesian would do.
\ref{tab:regression} shows that the slope for bad news for the
simulation\footnote{The details on how the data was simulated are explained in the Appendix}
of the Bayesian updates is much flatter than the slope in the data.

@@

@@

\hypertarget{model-fit}{%
\subsection{Model fit}\label{model-fit}}

@@

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

@@

@@

\hypertarget{experiment-instructions}{%
\section{Experiment Instructions}\label{experiment-instructions}}

@@

\emph{the full data generating process is printed in the rear side of
the instructions and always available to the subjects the matrices are
labeled with ``if your/the other participant's score is'' depending on
the treatment}

@@

  \bibliography{references.bib}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Learning with misspecified models: the case of overestimation},
  pdfauthor={Jimena Galindo},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=2.5cm]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

@@

\title{Learning with misspecified models: the case of overestimation}
\author{Jimena Galindo}
\date{October 13, 2023}

\begin{document}
\maketitle
\begin{abstract}
I design a framework and laboratory experiment that allows for the
comparison of multiple theories of misspecified learning. I focus on a
framework with endogenous information and two complementary
fundamentals: an ego-relevant parameter and a state. Within this
framework I study three forces that can lead to misspecified beliefs:
dogmatism, learning traps and biased updating. I find that biased
updating is the main driver of misspecifications. In addition, I vary
the degree of ego-relevance of the parameter by introducing a stereotype
treatment. The data is consistent with biased updating in both cases but
for potentially different reasons: subjects attribute successes to their
own ability and failures to luck. In the stereotype treatment they
over-compensate for initial negative biases by over-attributing positive
signals to the ability of others. This translates into similar observed
choices but different dynamisc in beliefs.
\end{abstract}

\newpage

@@

A growing body of literature in economics explores how people develop
incorrect beliefs about fundamentals. Most of this research centers on
scenarios where agents passively observe the world, and incorrectly
integrate the received information into their
beliefs\footnote{See \citet{benjamin2019} for a review of the literature on errors of 
probabilistic thinking}. However, many real-world situations cast agents
as active participants in the generation of information. In these cases,
the information they acquire is influenced by their actions and their
subsequent behavior is in turn determined by how they incorporate the
information into their beliefs. This further affects the information
generating process.

Consider a student who needs to decide how much effort to put into
studying for an exam. Their decision will depend on two different
factors: their belief about their intrinsic ability, and their belief
about how difficult the exam will be. The outcome that they observe will
be affected by how much they decide to study. Imagine that the student
puts in a moderate amout of effort and gets a surprisingly good grade.
Did they get a good grade because they are smarter than they thought? Or
because the exam was easier than they had anticipated? Their future exam
preparation strategy will depend fundamentally on which line of
reasoning they take; in turn, their performance in the next exam will
further affect the outcome they observe. This type of feedback loop is
what I call and \emph{endogenous information 
setting} and is at the center of the forces that I study.

I compare a set of theories that model learning in an endogenous
information setting and which can rationalize the persistence of
misspecified beliefs. To do so, I first come up with a unifying
framework in which each of the theories makes different predictions. I
then implement this framework in a laboratory experiment. The framework
that I use features an agent who, in order to make optimal decisions,
needs to learn a parameter that regards their own characteristics (an
\emph{ego-relevant} parameter), as well as an exogenously determined
state---in the case of the student it would be their intrinsic ability
and the difficulty of the exam.

Misspecified beliefs about an ego-relevant parameter are often referred
to as overconfidence and have been documented by behavioral
science\footnote{\citet{kelley1980} 
provides a review of the psychology literature.} and
economics\footnote{See \citet{benjamin2019} 
for a review of the literature in economics.} in a variety of settings.
\citet{Oster2013} find that subjects who are at high risk of having
Huntington's disease overestimate their probability of being healthy and
make retirement decisions as if they were. \citet{Hoffman2020} show that
workers overestimate the quality of their match to their current
employment and are unlikely to look for other opportunities.
\citet{Camerer1999} find that entrepreneurs are overconfident about the
quality of their enterprise, which leads to excessive entry and early
exit from markets.

In all of these examples, holding an incorrect belief about a
fundamental leads to suboptimal choices with potentially high costs. In
spite of the abundance of evidence, the scope of the existing research
is limited in terms of the frameworks it considers. Most of the
experimental evidence documenting the bias is collected in settings
where subject are passive learners.\footnote{
\citet{Gotte2022} and \citet{Ozyilmaz2022} are exceptions that study settings with endogenous information processes.}
They observe a noisy signal and report their beliefs over subsequent
rounds.\footnote{\citet{Bracha2012} and 
\citet{Mobius2022} are some examples}. Although these studies provide
important insights, they are not flexible enough to incorporate the
richer theories that have been proposed more recently. In particular,
they do not allow the studies of endogenous learning or for learning
about multiple parameters at once.\footnote{\citet{Coutts2020} 
studies an environment with an ego-relevant parameter and an exogenous state but does not 
incorporate the endogenous information process.}

In my experiment, I move away from the standard framework to study a
richer set of underlying mechanisms. In particular, the interaction
between the two parameters, together with an endogenous information
process, give rise to three possible explanations for the persistence of
incorrect beliefs: dogmatism, the existence of learning traps and
misattribution bias.

Within the first explanation, even with perfect Bayesian updating the
agent may fall into learning
traps.\footnote{Conditions under which this happens are detailed in \citet{Hestermann2021}}
These traps are characterized by a combination of misspecified beliefs
and optimal actions which produce information that confirms the
misspecified belief---this is a \emph{self-defeating equilibrium}. Once
an agent falls into it, the belief will be stable and even with a
correctly specified model of learning, they will not be able to abandon
their misspecified beliefs. If, in addition the agent is dogmatic about
their belief, \citet{Heidhues2018} show that they will inevitably fall
into a trap and thus will be able to rationalize and sustein their
initial dogmatically overconfident (or underconfident)
belief.\footnote{\citet{Gotte2022} study the case of agents with dogmatic initial beliefs in a 
laboratory experiment.}

\citet{Ba2023} moves away from dogmatism and endows the agents with a
mechanism through which they can abandon misspecified beliefs; this
allows them to avoid falling into learning traps. To do so agents
perform Bayesian hypothesis tests to evaluate which is the more likely
parameter out of two possibilities. By doing this, she characterizes the
set of situations in which, even agents who consider alternative
paradigms, may become trapped.\footnote{A 
similar mechanism is proposed by \citet{Schwarstein2021} in a setting with persuasion.}

Lastly, misattribution bias is the more classical explanation which has
been widely studied in behavioral
science.\footnote{see \citet{kelley1980} for a review.} Agents who
suffer from misattribution bias will attribute successes to their own
ability---the ego-relevant parameter---and failures to bad luck---the
state. Under this model of learning, even an agent who initially has a
correct initial belief may become overconfident if they observe a
sequence of successes. In this case, the main driver of the bias is not
an initial misspecification or the presence of learning traps, it is the
updating procedure
itself.\footnote{A more general framework that can be used to model 
this bias has also been proposed by \citet{Brunnermeier2005} and empirically studied by \citet{Bracha2012}.}

These theories provide the main building blocks for a simplified
framework that can be directly implemented in a laboratory experiment.
The goal is primarily to identify the which of the 3 forces---dogmatism,
the presence of learning traps or missatribution bias---better explains
the observed behavior. To do so, I compare the behavior predicted by
each theory to the benchmark given by the fully Bayesian updating
procedure.

On a secondary level, I also aim to understand if the misspecification
is inherently linked to the ego-relevance of one of the parameters or if
it is a more general phenomenon. To do so I vary the degree of
ego-relevance by introducing a treatment in which subjects learn about
the ability of another participant. In this treatment, the participants
know only the gender and nationality of the other, which can induce
stereotypes---a different type of misspecification.\\
If correct learning about the parameters happens at higher rates in the
stereotype treatment, it would suggest that the bias is intrinsically
linked to the ego-relevance of the parameter. In contrast, if similar
biased behavior arises in both treatments, it is more likely that the
main driver of these types of misspecified beliefs is the updating
procedure itself or the endogenous information process.

I find that, although some agents do fall into learning traps, the
behavior of most subjects is better explained by misattribution bias. I
also find that misattribution is no more prevalent in the ego-relevant
than in the stereotype condition. This suggests that the main driver of
the misspecification is the updating proceduce. However, the underlying
mechanism by which the bias is generated may be different in both
treatments: While in the ego-relevant condition, the intuitive
explanation behind the bias is that the agent prefers to hold themselves
in high esteem, in the stereotype condition it seems to be driven by
some sort of self-censoring---when subjects realize that they
underestimated the ability of another participant based on their gender
and nationality, they over-compensate by overestimating their ability.

Finally, I structurally estimate the parameters of the models to study
heterogeneity in the sample. I find that even at an individual level the
behavior is better explained by misattribution bias for most subjects.
There is a smaller group of subjects that can be better explained
through dogmatic beliefs and hypothesis testing and none of them behave
in line with the fully Bayesian benchmark.

In what follows I first discuss the theoretical framework and the
predictions of each of the theories. Then I introduce a unifying example
and my hypotheses. I then describe the experimental design and the
results, as well as the estimation of the parameters.

@@

\emph{The student must choose an effort level \(a\), which determines
her grade. For whatever the chosen effort level, the student must pay a
cost \(c(a) = \frac{1}{2}a^2\) and observes a grade of
\((\theta^*+a)\omega^*\). She repeats this process for infinitely many
periods. Assume also that the student's prior is such that
\(\mathbb{E}[\omega]= \omega^*\) and she is dogmatic about being of type
\(\hat{\theta}>\theta^*\).\footnote{The example is illustrated for an overconfident student but 
the results are symmetric for a dogmatic student who initially 
places probability one on some $\tilde{\theta}<\theta^*$.} Therefore,
the student's payoff in period \(t\) is given by }

@@

\emph{Under this specification, the myopic optimal effort level is
\(a_t^* = \omega^*\). Because the student does not know \(\omega^*\), he
will choose \(a_t = \mathbb{E}_t(\omega)\) where the expectation is
taken with respect to the student's belief at the beginning of period
\(t\). If she does not revise his effort choice for \(k\) periods, she
will receive an average utility of
\((\theta^*+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}\) but he was expecting
an average utility of
\((\hat{\theta}+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}\). In response, she
will apply Bayes rule to update his beliefs about \(\omega\) to get the
posterior belief with
\(\mathbb{E}_{t+k}[\omega] = \frac{(\theta^{*} + \omega^{*})\omega^{*}}{\hat{\theta} + \omega^{*}}\)
which is lower than the initial belief. This will cause the student to
choose a lower effort at \(t+k\). As a result, she will again receive an
average utility that is lower than what he expected which will cause her
belief to drift further down. This process will continue until the
average utility equals her expected utility under the dogmatic belief
that assigns probability 1 to \(\hat{\theta}\). At that point, the
student will have reached a self-defeating equilibrium and she will
continue to choose sub-optimal effort forever. }

@@

A regression analysis of this same data reveals that indeed, the slopes
are different for godd and bad news and that the difference is
statistically
significant.\footnote{I run a robustness check where I define
good news to be a net-positive sample (half or more of the outcomes observed in the sample received in that round
were successful) and bad news to be net-negative sample. I find the same pattern in this alternative regression model
and the results are reported in the appendix.} The regression results
are reported in Table \ref{tab:regression}. columns 2 and 3 of the table
consider the model for each treatment separately. I find no significant
differences in the parameters across treatment. This is consistent with
the fact that the bias is not driven by the ego-relevance of
\(\theta\).\footnote{The ego-relevance of $\theta$ is the motivation behind the self-attribution bias, 
however, the bias itself is due to an incorrect updating procedure in response to good and bad news. The presence of 
asymmetric updating in the stereotype condition can be due to self-censoring of the subjects. In particular, subjects 
who initially underestimated the performance of the other participant might have realized their mistake and 
by overestimating it after good news. This is consistent with the data presented here as well as with 
the data observed in the transition matrix for stereotypes which is reported in the Appendix}

@@

More convincing evidence against the Switching model is the lack of
inverse or attenuated correlation between the change in effort and the
news differential. In particular, if the switcher were the underlying
model, I would expect overconfident subjects who start with an
overestimation of \(\theta\) to increase their effort in response to
surprising bad news. This would flatten the slope of the regression line
for bad news relative to what a Bayesian would do. \ref{tab:regression}
shows that the slope for bad news for the
simulation\footnote{The details on how the data was simulated are explained in the Appendix}
of the Bayesian updates is much flatter than the slope in the data.

@@

The switching threshold is the parameter that determines the likelihood
ratio at which the switcher abandons the status quo paradigm and adopts
the alternative. In order to estimate this parameter, I assume that all
subjects are switchers and look at the rounds in which they chose to
reveal a probability matrix that was different from the one they
revealed in the previous round. I then calculate the likelihood ratio
for the alternative paradigm to the status quo. I average across all
such likelihood ratios and use the average as an estimate of the
switching threshold.

In detail, Let \(b_{t}^\theta\) be the belief about \(\theta\) in period
\(t\).\footnote{in the experimental data, 
I use the revealed matrix as a proxy for the belief. I take the matrix revealed in the previous period 
to be the status-quo and the matrix revealed in period $t$ to be the alternative}.
And consider a subject for whom
\(b_{\tau-1}^\theta \neq b_{\tau}^\theta\), where \(b_{t}^\theta\). That
is, in period \(\tau-1\) they chose to reveal a matrix that was
different from the one they revealed in period \(\tau\). The matrix
revealed in period \(\tau-1\) is the status quo paradigm and the matrix
revealed in period \(\tau\) is the alternative paradigm. The likelihood
ratio for the alternative paradigm to the status quo is \[
\ell = \frac{L(\theta | b_{\tau}^\theta)}{L(\theta | b_{\tau-1}^\theta)}=\frac{p_\tau(o_t|b_{\tau}^\theta)}{p_{\tau}(o_t|b_{\tau-1}^\theta)} 
\]

where \(p_\tau(o_t|b_{\tau}^\theta)\) is the probability of observing
the signal \(o_t\) when the agent is of type \(b_{\tau}^\theta\).

The estimate of the switching threshold, \(\alpha\), is the average of
the likelihood ratios across all subjects who switched in the
experiment, which is given by

\[\hat{\alpha} = \frac{1}{N}\sum_{i=1}^N \ell_{i\tau}\]

Where \(N\) is the number of switches observed in the experiment.

Because the feedback that subjects receive is the equivalent of 10
signal realizations, the estimate of the parameter is an upper bound of
the true parameter. A lower bound would be given by the average
likelihood ratio in the round before the switch happens. I choose to
estimate the upper bound to impose a harsher test on the theory.

Because the likelihood ratio is a function of the signal, it depends on
the sequence of signals that subjects observe. This means that in some
cases, the likelihood ratio increases by a large amount from one period
to another. Although such cases are not
common\footnote{In the data, only 1.8% of the 
observations have a likelihood ratio that exceeds 5.}, a few very large
values can skew the average. To avoid a biased parameter, I restrict my
sample to those observations for which the likelihood ratio is less than
5.\footnote{The full distribution of likelihood ratios is reported in the appendix}
Under these assumptions, the estimate of the switching threshold is
1.09.

This means that the evidence supporting the alternative paradigm needs
to be approximately 10\% more compelling than the evidence supporting
the status quo paradigm in order for the agent to switch. The estimation
treatment by treatment is slightly larger for the Ego condition than for
the Stereotype condition. However, the difference is not statistically
significant (p-value=0.96).

I then use this estimate to simulate the behavior of the switcher to
assess the fit of the model. Which is discussed in the next section.

@@

Estimation of the parameters of the self-attribution bias is more
involved than estimation of the switching threshold. In this case, there
is no obvious identification strategy that provides a simple estimate of
the parameters. Instead, I use the simulated method of moments to
estimate the parameters of the model.

In particular, I simulate the behavior of the self-attribution bias for
each possible prior and over a grid of parameters. I make the
simplifying assumption that the distortion is symmetric across good and
bad news. That is, I assume that
\(c(\theta_H, \omega, \text{good news})=c(\theta, \omega_L, \text{bad news})=:c_g\)
and that
\(c(\theta_M, \omega, \text{bad news})=c(\theta, \omega_M, \text{good news})=:c_b\).
In addition, I normalize the remaining cases to be equal to 1. This
leaves me with only 2 parameters to estimate.

The moments that I match in the data are the reaction to good news and
the reaction to bad news. In particular, I match the average change in
effort in response to good news and the average change in effort in
response to bad news. That is I match the following moments:

\[
\bar{G} = \frac{1}{N}\sum_{i=1}^N \frac{1}{T}\sum_{t=1}^T \mathbb{I}\{o_t > \bar{o}\} \cdot (e_t - e_{t-1}) 
\] \[
\bar{B} = \frac{1}{N}\sum_{i=1}^N \frac{1}{T}\sum_{t=1}^T \mathbb{I}\{o_t < \bar{o}\} \cdot (e_t - e_{t-1})
\] where \(N\) is the number of subjects in the sample, \(T\) is the
number of rounds in the experiment.

I then use the simulated method of moments to estimate the parameters of
the model. In particular, I simulate the behavior of the model for each
possible prior and over a grid of parameters. I then calculate the
average change in effort in response to good news and the average change
in effort in response to bad news for each possible prior and over the
grid of parameters and across \(I\) simulated indiviaduals. I then
minimize the objective function given by the sum of squared differences
between the simulated moments and the empirical moments. That is, I
solve the following problem

\hypertarget{model-fit}{%
\section{Model fit}\label{model-fit}}

@@

\newpage
\appendix

\hypertarget{appendix}{%
\section{appendix}\label{appendix}}

@@

\hypertarget{full-distribution-of-likelihood-ratios}{%
\subsection{Full distribution of likelihood
ratios}\label{full-distribution-of-likelihood-ratios}}

The likelihood ratios at rounds in which I observe a change in the
belief about \(\theta\) presents a long right-tail. In particular, 1.8\%
of the observations have a likelihood ratio that exceeds 5 which would
mean that the agent requires evidence that is 5 times more compelling in
favor of the alternative paradigm in order to switch. Figure
\ref{fig:lr-hist} shows the distribution of likelihood ratios for all
observations.

\begin{figure}
\hypertarget{fig:lr-hist}{%
\centering
\includegraphics{../figures/lr_hist.png}
\caption{Distribution of likelihood ratios for all observed belief
changes}\label{fig:lr-hist}
}
\end{figure}

@@

\hypertarget{reaction-to-good-and-bad-news-robustness-check}{%
\subsection{Reaction to good and bad news (robustness
check)}\label{reaction-to-good-and-bad-news-robustness-check}}

I use an alternate definition for good and bad news and check whether
the results are robust to this. I define a signal to be a net positive
if the number of successes is equal to or greater than the number of
failures. I define a signal to be a net negative if the number of
failures is greater than the number of successes. I now treat net
positive signals as good news and net negative signals as bad news. I
then rerun the regression analysis and find that the results are robust
to this alternative definition. In particular, net negative signals are
associated with a stark decrease in effort for both treatments. Net
positive signals are associated with a small decrease in effort for both
treatments. The results are reported in Table \ref{tab:regression-net}.

\begin{table}[!htbp] \centering 
  \caption{Regression of effort change on news difference with robust standard errors} 
  \label{tab:regression-net} 
\begin{tabular}{@{\extracolsep{5pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:}} \\ 
\cline{2-4} 
\\[-1.8ex] & \multicolumn{3}{c}{Change in effort} \\ 
 & All & Ego-relevant & Stereotype \\ 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\hline \\[-1.8ex] 
 Net-positive signal & 0.99$^{***}$ & 0.80$^{***}$ & 1.24$^{***}$ \\ 
  & (0.05) & (0.05) & (0.05) \\ 
  & & & \\ 
 Signal value & 0.21$^{***}$ & 0.22$^{***}$ & 0.21$^{***}$ \\ 
  & (0.02) & (0.02) & (0.02) \\ 
  & & & \\ 
 Signal value * Net-Positive signal & $-$0.28$^{***}$ & $-$0.26$^{***}$ & $-$0.31$^{***}$ \\ 
  & (0.02) & (0.02) & (0.02) \\ 
  & & & \\ 
 Constant & $-$0.49$^{***}$ & $-$0.48$^{***}$ & $-$0.50$^{***}$ \\ 
  & (0.04) & (0.04) & (0.04) \\ 
  & & & \\ 
\hline \\[-1.8ex] 
Observations & 4,680 & 2,700 & 1,980 \\ 
R$^{2}$ & 0.05 & 0.05 & 0.05 \\ 
Adjusted R$^{2}$ & 0.05 & 0.04 & 0.05 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\hypertarget{experimental-instructions}{%
\section{Experimental Instructions}\label{experimental-instructions}}

@@

\emph{The full data generating process is printed on the rear side of
the instructions and always available to the subjects throughout part
2.} \emph{Each of the matrices is labeled with ``if your/the other
participant's score is\ldots{}'' depending on the treatment}

@@

\newpage

\renewcommand\refname{References}
  \bibliography{references.bib}
